Assessment of forestry road network projection methods using the “roads” package in R

### 2022-08-26 analysis/paper/paper.Rmd
change high density #2 and #4

#set paths
data_path_raw <- "analysis/data/raw_data/"
data_path_drvd <- "analysis/data/derived_data/"

@@

We projected road development in the Revelstoke TSA from 1990 to the present using several different inputs and compared the projections to the observed present day road network in order to evaluate the accuracy of the projections and their sensitivity to different inputs. We compared two resolutions for the cost surface by aggregating the fine resolution cost raster to a coarser 1000 m resolution. We also compared methods for locating landing points within cutblocks by comparing using the polygon centroid, regularly spaced and random points, as well as point densities of 1 and 100 pts/km^2^ (Figure \@ref(fig:fig2)). Note that the default skidding distance used by the Hardy QGIS plugin is 100 m which corresponds to 25 pts/km^2^ .

@@

When comparing between projection methods we found that the density of landing points chosen within cutblocks made a large difference to the results with the high density method overestimating the road density and road presence metrics while the low density and centroid methods underestimated them. The sampling method for choosing points in the cutblocks, either random or regular, did not have a large effect on the results for the low sample density but regular sampling was less accurate for all metrics with the high sample density. Hardy's QGIS plugin performed well at the overall scale but tended to overestimate the amount of roads within cutblocks.

@@

```{r fig5, fig.height=9, fig.width=fig_widths["two"], fig.cap="Comparison of performance metrics for each projection method. F measure is the harmonic mean of precision and sensitivity, precision is the proportion of  predicted roads that are observed roads and sensitivity is the proportion of observed roads that were correctly predicted."}

matchData_sum <- matchData %>% ungroup() %>%
  mutate(agreement = factor(agreement,
                            levels = c("False negative", "False positive",
                                       "Agree roaded","Agree roadless",
                                       "Pre-existing roads")),
         sampleDens = paste(sampleType, sampleDens) %>%
           str_replace("1e-06", "low density") %>%
           str_replace("1e-04", "high density") %>%
           str_replace("centroid high density", "centroid") %>%
           str_replace("klementQGIS NA", "klementQGIS") %>%
           factor(levels = c("centroid",
                             "random low density",
                             "regular low density",
                             "random high density",
                             "regular high density",
                             "klementQGIS"),
                  labels = c("Centroid", 
                             "Random\nlow density",
                             "Regular\nlow density",
                             "Random\nhigh density",
                             "Regular\nhigh density",
                             "Hardy QGIS")),
         metric = factor(metric,
                         levels = c("roadPresence",
                                    "forestryDisturbance",
                                    "roadDisturbance"),
                         labels = c("Road presence",
                                    "Forestry disturbance\nfootprint",
                                    "Road disturbance footprint"))) %>%
  group_by(metric, sampleDens, agreement) %>%
  summarise(count = sum(count), .groups = "drop_last") %>%
  mutate(perc = count/sum(count)*100)

>>>>

#set paths
data_path_raw <- "analysis/data/raw_data/"
data_path_drvd <- "analysis/data/derived_data/TSA27"

@@

We projected road development in the Revelstoke TSA from 1990 to the present using several different inputs and compared the projections to the observed present day road network in order to evaluate the accuracy of the projections and their sensitivity to different inputs. We compared two resolutions for the cost surface by aggregating the fine resolution cost raster to a coarser 1000 m resolution. We also compared methods for locating landing points within cutblocks by comparing using the polygon centroid, regularly spaced and random points, as well as point densities of 1 and 10 pts/km^2^ (Figure \@ref(fig:fig2)). Note that the default skidding distance used by the Hardy QGIS plugin is 100 m which corresponds to 25 pts/km^2^ .

@@

When comparing between projection methods we found that the density of landing points chosen within cutblocks affected the results, with the high density method underestimating the road density and road presence metrics less than the low density and centroid methods. The sampling method for choosing points in the cutblocks, either random or regular, did not have a large effect on the results. Hardy's QGIS plugin performed well at the overall scale but tended to overestimate the amount of roads within cutblocks.

@@

```{r fig5, fig.height=6, fig.width=fig_widths["mid"], fig.cap="Comparison of performance metrics for each projection method. F measure is the harmonic mean of precision and sensitivity, precision is the proportion of  predicted roads that are observed roads and sensitivity is the proportion of observed roads that were correctly predicted."}

matchData_sum <- matchData %>% ungroup() %>%
  mutate(agreement = factor(agreement,
                            levels = c("False negative", "False positive",
                                       "Agree roaded","Agree roadless",
                                       "Pre-existing roads")),
         sampleDens = paste(sampleType, sampleDens) %>%
           str_replace("1e-06", "low density") %>%
           str_replace("1e-05", "high density") %>%
           str_replace("centroid low density", "centroid") %>%
           str_replace("klementQGIS NA", "klementQGIS") %>%
           factor(levels = c("centroid",
                             "random low density",
                             "regular low density",
                             "random high density",
                             "regular high density",
                             "klementQGIS"),
                  labels = c("Centroid", 
                             "Random\nlow density",
                             "Regular\nlow density",
                             "Random\nhigh density",
                             "Regular\nhigh density",
                             "Hardy QGIS")),
         metric = factor(metric,
                         levels = c("roadPresence",
                                    "forestryDisturbance",
                                    "roadDisturbance"),
                         labels = c("Road presence",
                                    "Forestry disturbance\nfootprint",
                                    "Road disturbance footprint"))) %>%
  group_by(metric, sampleDens, agreement) %>%
  summarise(count = sum(count), .groups = "drop_last") %>%
  mutate(perc = count/sum(count)*100)
